Model based reinforcement learning at FERMI FEL

### 2020-11-16 main.tex
txs auto checkin

\section{Deep Reinforcement learning}
Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\mathcal{R}$ an initial state distribution $d_0$.
Goal of reinforcement learning is to find a $\pi(s)\mapsto a$, which is the solution of:
\begin{align}
\max J(\pi)  =  \mathbb E_{\tau \thicksim p_{\pi}}\left[\sum_{t=0}^{H}\gamma^t r(s_t,a_t)\right],
\end{align}
where $p_\pi = d_0(s_0)\prod_{t_0}^H\pi(a_t,s_t)\mathcal T(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g. a deep neural network. In the first case one speaks about policy gradient methods in the latter about approximate dynamic programming, which we now discuss.
\subsection{Approximate dynamic programming}
The state-value function $Q(s,a)$ is expressed as $Q_\phi(s,a)$, where $\phi$ denotes the parameters of the function approximator. By satisfying the Bellmann-optimality operator the optimal $Q^*(s,a)$ can be found:
\begin{align}
    \min_\theta \left(\vec Q_\theta - \mathcal{B}^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
\end{align}
This operator has a unique fixed point but is non-linear, due to the $\max$ - operator:
\begin{align}
 \mathcal{B}^*\vec Q_\theta(s_t,a_t) = r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
\end{align}
The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{}.
One way to reduce the effect is to take a simple analytical form of the $Q$-function.
\subsection{NAF}
If a specific quadratic form of the $Q$ function is assumed:
\begin{align}
     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s).
\end{align}
 The optimal policy is given by:
\begin{align}
    \pi^*(s)=(\delta(a)-\arg\!\max_a Q_\theta(s,a)).
\end{align}
One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
The maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
\begin{align}
	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta\prime,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
\end{align}
$\theta\prime$ denotes the weights of a previous iteration. To stabilize the network training a small artificial noise is added to the actions in the update.
%\subsection{New NAF network architecture}

@@

	\item The MFRL agent as e.g. the TRPO and PPO or the SAC.

@@

The number of data-points taken every time the model as improved was firstly determined by the number of initial data points, collect using a random policy. The initialization phase has to be chosen not to small to minimize the chance of getting trapped in a local minimum for too long. Afterwards at least on full episode should be taken. We decided to use a short total episode length to reduce the impact of the compound error, the accumulation error following a wrong model. The maximal number of steps was ten. During the data collected on the real system, the latest policy was taken with some small noise added to improve the exploration.\\

>>>>

\section{Deep Reinforcement learning}
Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
Goal of reinforcement learning is to find a $\policy\mapsto a$, which is the solution of:
\begin{align}
\max J (\policy)  =  \mathbb E_{\traj \thicksim p_{\olicyi}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],
\end{align}
where $p_\pi = d_0(s_0)\prod_{t_0}^H\pi(a_t,s_t)\mathcal T(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g. a deep neural network. In the first case one speaks about policy gradient methods in the latter about approximate dynamic programming, which we now discuss.
\subsection{Approximate dynamic programming}
The state-value function $Q(s,a)$ is expressed as $Q_\phi(s,a)$, where $\phi$ denotes the parameters of the function approximator. By satisfying the Bellmann-optimality operator the optimal $Q^*(s,a)$ can be found:
\begin{align}
    \min_\theta \left(\vec Q_\theta - \bellman^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
\end{align}
This operator has a unique fixed point but is non-linear, due to the $\max$ - operator:
\begin{align}
 \bellman^*\vec Q_\theta(s_t,a_t) := r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
\end{align}
The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{}.
One way to reduce the effect is to take a simple analytical form of the $Q$-function.
\subsection{NAF}
If a specific quadratic form of the $Q$ function is assumed:
\begin{align}
     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s).
\end{align}
 The optimal policy is given by:
\begin{align}
    \pi^*(s)=(\delta(a)-\arg\!\max_a Q_\theta(s,a)).
\end{align}
One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
The maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
\begin{align}
	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta\prime,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
\end{align}
$\theta\prime$ denotes the weights of a previous iteration. To stabilize the network training a small artificial noise is added to the actions in the update. This algorithm has an extremely good sample-efficiency for suitable problems as can be found in accelerators.
%\subsection{New NAF network architecture}

@@

	\item The MFRL agent as e.g. the TRPO and PPO or the SAC and its training

@@

The number of data-points taken every time the model as improved was firstly determined by the number of initial data points, collect using a random policy. The initialization phase has to be chosen not to small to minimize the chance of getting trapped in a local minimum for too long. Afterwards at least on full episode should be taken. We decided to use a short total episode length to reduce the impact of the compound error, the accumulation error following a wrong model, as well known for Dyna-style approaches. The maximal number of steps was ten. During the data collected on the real system, the latest policy was taken with some small noise added to improve the exploration.\\
